---
title: 'NEW YORK AIRBNB - DATA VISUALIZATIONS'
author: "Cindy Chen"
date: "3/19/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggmap)
library(tmap)
library(ggplot2)
library(readr)
library(ggthemes)
library(DT)
library(RJSONIO)
library(stringr)
library(R.utils)
library(geojsonio)
library(rgdal)
library(broom)
library(ggrepel)
library(viridis)
library(RColorBrewer)
library(leaflet)
```

## Question 1: Overall Location

a) Below is a map of where in NYC AirBnB listings are located
```{r nyc_airbnb}
#unzip the zip folder (which I renamed by removing the .csv name at the end)
#unzip("data/airbnb_listings.zip")

#load airbnb data from unzipped folder
airbnb_data <- read_csv("airbnb_listings.csv", show_col_types = FALSE)

#let's look at the data frame to better understand it
head(names(airbnb_data)) #I've set this to head() to avoid printing out a long list
```

```{r createmap1}
nyc_map <- geojson_read("data/neighbourhoods.geojson", what = "sp") %>% tidy()

ggplot() +
  geom_polygon(data =nyc_map, aes(x = long, y = lat, group = group), fill="#d3d3d3", color = "white") +
  geom_point(data = airbnb_data,
             aes(x = longitude, y = latitude), color = "coral",
             size=0.1,
             alpha = 0.2) +
  labs(title = "AirBnB's in New York City: \nMost Populous in Manhattan and West Brooklyn") +
  theme_map() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())

```

```{r}
nyc_heat_map <- ggplot() +
  geom_polygon(data =nyc_map, aes(x = long, y = lat, group = group), fill="#d3d3d3", color = "white") +
  theme_map() +
  stat_density2d(data = airbnb_data,
                 geom = "polygon",
                 aes(x=longitude,y=latitude,
                     fill = ..level..,
                     alpha = ..level..)) +
  scale_fill_gradient(low = "yellow", high = "red",
                      name = "AirBnb listing density") +
  scale_alpha(guide = "none") +
  #add Hell's Kitchen
  annotate("text",x=-73.98355, y = 40.77236, label="Hell's Kitchen",
                   color="#483d8b", fontface = 2, size=2) +
  #add Williamsburg
  annotate("text",x=-73.9256, y = 40.72341, label="Williamsburg",
                   color="#483d8b", fontface = 2, size=2) +
  #add East Village
  annotate("text",x=-73.97543, y =40.73396, label="East Village",
                   color="#483d8b", fontface = 2, size=2) +
  labs(title = "Density of NYC AirBnb listings:\nConcentrated in Hell's Kitchen, the East Village,\n& Williamsburg")

nyc_heat_map +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(size = 12))
  
```

## Question 2. Renting out your apartment vs. permanent rentals

a) Year-round vs sporadic rentals

_If a listing is available on AirBnb for at least 75% of the year, I use this threshold to identify permanent listings.  For sporadic listings, I set the threshold to less than 20% AND more than 1 day of availability.  Anything in-between, I consider a semi-permanent listing._

I created two (2) sets of maps for each question, one using tmap and the other using ggmap.

```{r pressure, echo=FALSE}
#understand the different values under the column availability_365
str(airbnb_data$availability_365)

#create a data frame where listings are classified as permanent or sporadic based on their
#availability
permanent_vs_spor <- airbnb_data %>%
  select(latitude, longitude, availability_365, neighbourhood_cleansed, neighbourhood_group_cleansed) %>%
  mutate(long_term_flag = case_when(availability_365/365 >= 0.75 ~ "PERMANENT",
                                    (availability_365/365 < 0.2 & availability_365 > 0) ~ "SPORADIC",
                                    TRUE ~ "SEMI-PERMANENT"))
```


```{r map_3}
ggplot() +
  geom_polygon(data =nyc_map, aes(x = long, y = lat, group = group), fill="#d3d3d3", color = "white") +
  theme_map() +
  geom_point(data = filter(permanent_vs_spor, long_term_flag == "PERMANENT"),
             aes(x = longitude, y = latitude), color = "coral",
             size=0.1,
             alpha = 0.2) +
  labs(title = "Year-Round (Permanent) AirBnB's in New York City\nTop 5 Neighbourhoods Highlighted") +
  geom_text_repel() +
  #add Hell's Kitchen
  annotate("text",x=-74.01235, y = 40.77236, label="Hell's Kitchen",
                   color="#483d8b", fontface = 2, size=2) +
  #add Williamsburg
  annotate("text",x=-73.9256, y = 40.72341, label="Williamsburg",
                   color="#483d8b", fontface = 2, size=2) +
  #add BedStuy
  annotate("text",x=-73.90606, y =40.70021, label="Bed-Stuy",
                   color="#483d8b", fontface = 2, size=2) +
  #add Midtown
  annotate("text",x=-73.96088, y = 40.76811, label="Midtown",
                   color="#483d8b", fontface = 2, size=2) +
  #add Harlem
  annotate("text",x=-73.95905, y = 40.79724, label="Harlem",
                   color="#483d8b", fontface = 2, size=2)
```
```{r}
#create a cleaner data set that we will use to merge into nyc_map_raw
mini_permanent_spor <- permanent_vs_spor %>%
  group_by(neighbourhood_cleansed, long_term_flag) %>%
  summarize(count_of_listings = n())
nyc_map_raw <- geojson_read("data/neighbourhoods.geojson", what = "sp") 

combined <- nyc_map_raw@data %>% 
  left_join(filter(mini_permanent_spor, long_term_flag == "PERMANENT"),
            by = c(neighbourhood = "neighbourhood_cleansed"))

nyc_map_raw@data <- combined #create new data set that has the permanent vs semi-permanent flags

hotspots <- ifelse(is.na(nyc_map_raw$count_of_listings > 400) == TRUE, FALSE, nyc_map_raw$count_of_listings > 400)

tm_shape(nyc_map_raw) +
  tmap_options(check.and.fix = TRUE) +
  tm_fill("count_of_listings",
          title = "Permanent AirBnb Listings by Neighbourhood",
          convert2density = TRUE,
          style = "kmeans") +
  tm_borders(alpha = .5) +
  tm_shape(nyc_map_raw[hotspots,]) +
  tm_text("neighbourhood", size = .6) 
```

```{r map_4}
ggplot() +
  geom_polygon(data =nyc_map, aes(x = long, y = lat, group = group), fill="#d3d3d3", color = "white") +
  theme_map() +
  geom_point(data = filter(permanent_vs_spor, long_term_flag == "SEMI-PERMANENT"),
             aes(x = longitude, y = latitude), color = "sky blue",
             size=0.1,
             alpha = 0.1) +
  labs(title = "Semi-Permanent AirBnB's in New York City:\nTop 5 Neighbourhoods Highlighted") +
  #add Upper West Side
  annotate("text",x=-73.99381, y = 40.78798, label="Upper West Side",
                   color="#483d8b", fontface = 2, size=2) +
  #add Williamsburg
  annotate("text",x=-73.9256, y = 40.72341, label="Williamsburg",
                   color="#483d8b", fontface = 2, size=2) +
  #add BedStuy
  annotate("text",x=-73.90606, y =40.70021, label="Bed-Stuy",
                   color="#483d8b", fontface = 2, size=2) +
  #add Bushwick
  annotate("text",x=-73.94098, y = 40.68116, label="Bushwick",
                   color="#483d8b", fontface = 2, size=2) +
  #add Harlem
  annotate("text",x=-73.95905, y = 40.79724, label="Harlem",
                   color="#483d8b", fontface = 2, size=2)
```

```{r semiperm_dnesity}
#reload the raw spatial data since we would have appended data to it
nyc_map_raw <- geojson_read("data/neighbourhoods.geojson", what = "sp") 

#now add semi-permanent listing data
combined_2 <- nyc_map_raw@data %>% 
  left_join(filter(mini_permanent_spor, long_term_flag == "SEMI-PERMANENT"),
            by = c(neighbourhood = "neighbourhood_cleansed"))

nyc_map_raw@data <- combined_2 #create new data set that has the permanent vs semi-permanent flags

hotspots_semi <- ifelse(is.na(nyc_map_raw$count_of_listings > 900) == TRUE, FALSE, nyc_map_raw$count_of_listings > 900)

tm_shape(nyc_map_raw) +
  tmap_options(check.and.fix = TRUE) +
  tm_fill("count_of_listings",
          title = "Count of Semi-Permanent AirBnb Listings",
          convert2density = TRUE,
          style = "kmeans") +
  tm_borders(alpha = .5) +
  tm_shape(nyc_map_raw[hotspots_semi,]) +
  tm_text("neighbourhood", size = .6) 
```

```{r}
#locate neighborhoods with most permanent rentals, rank them, and visualize
permanent_vs_spor %>%
  group_by(neighbourhood_cleansed, long_term_flag, neighbourhood_group_cleansed) %>%
  count() %>%
  filter(long_term_flag == "PERMANENT") %>%
  arrange(desc(n)) %>%
  rename(new_count = n) %>%
  filter(new_count > 200) %>%
  ggplot() +
  geom_col(aes(x = reorder(neighbourhood_cleansed, new_count), y = new_count, fill = neighbourhood_group_cleansed)) +
  theme_classic() +
  coord_flip() +
  theme(legend.position = "bottom") +
  labs(x = "Neighbourhood",
       y = "Number of AirBnbs",
       title = "Top Permanent AirBnb Counts by NYC Neighbourhood") +
  scale_fill_manual(name="Borough",
                      values = c("#481567ff", "#B8DE29FF")) +
  theme(plot.title = element_text(size = 11))
```
```{r}
#locate neighborhoods with most permanent and semi-permanent rentals, rank them, and visualize the split of permanent vs semi-permanent
permanent_vs_spor %>%
  group_by(neighbourhood_cleansed, long_term_flag, neighbourhood_group_cleansed) %>%
  count() %>%
  arrange(desc(n)) %>%
  rename(new_count = n) %>%
  filter(long_term_flag != "SPORADIC") %>%
  group_by(neighbourhood_cleansed) %>%
  mutate(total_listings = sum(new_count)) %>%
  ungroup() %>%
  filter(total_listings > 800) %>%
  ggplot() +
  geom_col(aes(x = reorder(neighbourhood_cleansed, total_listings), y = new_count, fill = long_term_flag)) +
  scale_fill_viridis_d() +
  theme_classic() +
  coord_flip() +
  theme(legend.position = "bottom") +
  labs(x = "Neighbourhood",
       y = "Number of AirBnbs",
       title = "Top Permanent & Semi-Permanent AirBnb Counts by NYC Neighbourhood") +
  theme(plot.title = element_text(size = 11))
```

b) Data table of the top hosts, total listing count, avg nightly price, and estimated avg income


_To calculate estimated average monthly total income, I multiplied by 30 days as a broad assumption if they were to rent their listing for 30 days a month._
```{r multiple_rentals}
#determine the top hosts; I used this method to double check that the listing count matched
top_hosts <- airbnb_data %>%
  group_by(host_id) %>%
  summarize(count_listings = n_distinct(listing_url)) %>%
  arrange(desc(count_listings)) %>%
  slice_head(n = 20)

avg_price_income <- airbnb_data %>%
  #convert price to numeric
  mutate(convert_price = gsub(",", "", price)) %>%
  mutate(convert_price = as.numeric(str_remove(convert_price, "[$]"))) %>%
  mutate(monthly_income = 30 * convert_price) %>%
  group_by(host_id) %>%
  summarize(avg_price = round(mean(convert_price),2),
            avg_monthly_income = round(mean(monthly_income * host_listings_count),2))
```

```{r}
#join data frames
top_host_income <- top_hosts %>%
  left_join(avg_price_income) %>%
  rename(AirBnb_Host_ID = host_id,
         Total_Listings = count_listings,
         Avg_Price = avg_price,
         Avg_Monthly_Income = avg_monthly_income)

#create pretty headers
pretty_headers <- gsub("[_]", " ", colnames(top_host_income))

#create data table
datatable(top_host_income,
          filter = list(position = "top"),
            colnames = pretty_headers) %>%
  formatStyle('Avg_Price',
              background = styleColorBar(range(top_host_income$Avg_Price),'lightblue'),
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center') %>%
  formatCurrency('Avg_Price',
                 currency = "$",
                 digits = 2,
                 mark = ",") %>%
  formatCurrency('Avg_Monthly_Income',
                 currency = "$",
                 digits = 0,
                 mark = ",") %>%
  formatStyle('Avg_Price',
              background = styleColorBar(range(top_host_income$Avg_Price),'lightblue'),
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center') %>%
  formatStyle('Avg_Monthly_Income',
              background = styleColorBar(range(top_host_income$Avg_Monthly_Income),'gold'),
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center') 

```

## Question 3. Top Reviewed Rentals

```{r}
#prepare data
top_rated <- airbnb_data %>%
  select(latitude, longitude, review_scores_rating, number_of_reviews, price, guests_included, room_type, property_type) %>%
  arrange(desc(review_scores_rating), desc(number_of_reviews)) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id <= 100) %>%
  mutate(top_what = "TOP RATED") #add a flag

most_expensive <- airbnb_data %>%
  select(latitude, longitude, price, guests_included, room_type, property_type) %>%
  mutate(convert_price = gsub(",", "", price)) %>%
  mutate(convert_price = as.numeric(str_remove(convert_price, "[$]"))) %>%
  arrange(desc(convert_price)) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id <= 100) %>%
  mutate(top_what = "MOST EXPENSIVE") #add a flag

top_rate_expensive <- top_rated %>%
  full_join(most_expensive)
```

```{r}
#set color palette
cindys_palette = colorFactor(c("#ff8a87", "#29c8e3"), domain = top_rate_expensive$top_what)
color_4_top  = cindys_palette(top_rate_expensive$top_what)

popup_content <- paste("<b>Property Type:</b>", top_rate_expensive$property_type, "<br/>",
                                 "<b>Guests Included:</b>", top_rate_expensive$guests_included, "<br/>",
                                 "<b>Room Type:</b>", top_rate_expensive$room_type, "<br/>",
                       "<b>Price:</b>", top_rate_expensive$price, "<br/>")

leaflet(top_rate_expensive) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.VoyagerLabelsUnder") %>%
  addCircles(lng = ~longitude, lat = ~latitude, color = color_4_top) %>%
  addLegend(pal = cindys_palette, values = ~top_rate_expensive$top_what, title = "Most Expensive or Top Rated") %>%
  addCircleMarkers(popup = popup_content, clusterOptions = markerClusterOptions()) %>%
  clearBounds() %>%
  addEasyButton(easyButton(
    icon = "fa-crosshairs",
    title = "Locate Me",
    onClick = JS("function(btn, map){ map.locate({setView: true}); }")
  ))
```

### Bonus Question: Subway Access & Airbnb

I only got through part of it but it was fun to experiment.
```{r}
#load data on NYC subways
nyc_subway_entrance <- readOGR(dsn = "data/nyc_subway_map/entrances_nyc_subway", layer = "subway_entrances_may2016")
nyc_subway_stns <- readOGR(dsn = "data/nyc_subway_map/stops_nyc_subway", layer = "stops_nyc_subway_jan2017")

#load back in the nyc_map
nyc_map_raw <- geojson_read("data/neighbourhoods.geojson", what = "sp") 
```
The neighbourhood I will focus on is Bed-Stuy.
```{r}
library(rgeos)
library(sf)

bed_stuy_airbnb <- airbnb_data %>%
  filter(neighbourhood_cleansed == "Bedford-Stuyvesant") %>%
  select(longitude, latitude, price, id)

#convert the airbnb data into a spatial object so that we can run calculations on it
bed_stuy <-SpatialPointsDataFrame(coords = bed_stuy_airbnb[,c(1,2)], data = bed_stuy_airbnb,
                               proj4string = CRS(st_crs(nyc_subway_entrance)$proj4string))

#filter onto areas framed by BedStuy
bed_stuy_only <- nyc_map_raw$neighbourhood == "Bedford-Stuyvesant"
```


```{r}

bed_stuy_sp <- bed_stuy_airbnb %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "NAD83") #will not work when I try to project it using the same CRS as the subway

#visualize BedStuy Airbnb's and subway stations
tm_shape(nyc_map_raw[bed_stuy_only, ]) +
  tmap_options(check.and.fix = TRUE) +
  tm_borders() +
  tm_shape(bed_stuy_sp) +
  tm_dots(col = "red") +
  tm_shape(nyc_subway_entrance) +
  tm_dots(col = "black", size = 0.8)
```


```{r}
library(tidyr)
#calculate the distances between 
calc_distances <- gDistance(bed_stuy, nyc_subway_entrance, byid=T)

#rename the rows and columns
rownames(calc_distances) <- nyc_subway_entrance$Station_Na
colnames(calc_distances) <- bed_stuy$id

closest_stn_airbnb <- calc_distances %>%
  as.data.frame() %>%
  pivot_longer(col = everything(),
               names_to = "airbnb",
               values_to = "distance") %>%
  mutate(station_name = rep(nyc_subway_entrance$Station_Na, 3708)) %>%
  group_by(airbnb) %>% #group by airbnb
  filter(distance == min(distance)) %>% # filter onto the rows with the closest station entrance for each airbnb
  mutate(airbnb = as.numeric(airbnb))

#this gives me the closest station for each AirBnB in BedStuy
head(closest_stn_airbnb)
```

```{r}
#convert this new data frame back into a spatial object
closest_stn_airbnb_2 <- closest_stn_airbnb %>%
  left_join(bed_stuy_airbnb, c("airbnb" = "id")) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "NAD83") #will not work when I try to project it using the same CRS as the subway

#closest_sp <- SpatialPointsDataFrame(coords = closest_stn_airbnb_2[,c(4, 5)], data = closest_stn_airbnb_2,
#                               proj4string = CRS(st_crs(nyc_subway_entrance)$proj4string))

#visualize BedStuy Airbnb's and subway stations
tm_shape(nyc_map_raw[bed_stuy_only, ]) +
  tm_borders() +
  tm_shape(closest_stn_airbnb_2) +
  tm_dots(col = "station_name", legend.show = FALSE) +
  tm_shape(nyc_subway_entrance) +
  tm_dots(col = "black", size = 1, legend.show = FALSE)

#this is incorrect but I could not figure out why.  I assume it maps my labels correctly when gDistance() is run, but I really think that is where the issue is.
```

