---
title: "Shiny App - Billboard Music Charts"
author: "Cindy Chen"
date: "4/14/2022"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(highcharter)
library(dplyr)
library(viridisLite)
library(dplyr)
library(shiny)
library(DT)
library(plotly)
library(stringr)
```

#### Overview

This assignment focuses on creating three individual Shiny apps to search and visualize top songs, top artists, and the relationship between audio features among songs on the top Billboard charts.

#### Data Prep

```{r}
billboard <- read.csv("billboard.csv")
audio_features <- read.csv("audio_features.csv")

billboard_dt <- billboard %>%
  mutate(year = as.integer(format(as.Date(week_id, format = "%m/%d/%Y"), "%Y"))) %>%
  group_by(song, performer, year) %>%
  summarize(weeks_on_top100 = max(weeks_on_chart),
            peak_position = min(peak_position)) %>%
  ungroup()

billboard_artists <- billboard %>%
  mutate(song_date = as.Date(week_id, format = "%m/%d/%Y")) %>%
  dplyr::select(performer, song, song_date, week_position) %>%
  unique()

#create pretty headers
pretty_headers <- gsub("[_]", " ", colnames(billboard_dt))

# left join spoitfy and billboard data
spotify_billboard <- billboard %>%
  left_join(audio_features, by = c("performer" = "performer", "song" = "song", "song_id" = "song_id")) %>%
  mutate(year = as.integer(format(as.Date(week_id, format = "%m/%d/%Y"), "%Y"))) %>%
  group_by(song, performer, spotify_genre, year, tempo, valence) %>%
  summarize(top_position = min(week_position)) %>%
  ungroup() %>%
  mutate(Top_Type = case_when(top_position <= 1 ~ 1,
                              top_position <= 10 ~ 10,
                              top_position <= 20 ~ 20)) %>%
  filter(Top_Type <= 20,
         year >= 2004 & year <= 2014,
         is.na(valence) == FALSE,
         is.na(tempo) == FALSE)

# To subset the data to make it more manageable, I chose songs that were part of the top charts and were between
# 2004 and 2014 (because that was when I was a teen!)
```


### Question 1. Top Songs

```{r topsongs_shiny}
### UI ###
ui <- fluidPage(
  title = "Top Songs",
  h2("Top Songs"),
  
    sliderInput(
      inputId = "choose_year", 
    label = "Choose the years",
    min = 1997,
    max = 2021,
    value = c(1997, 2021)),
  DT::dataTableOutput("topsong_table"))

### SERVER ###
server <- function(input, output) {
my_range1 <- reactive({input$choose_year[1]})
my_range2 <- reactive({input$choose_year[2]})

billboard_dt_new <- reactive({billboard_dt %>%
    filter(year >= my_range1() &
             year <= my_range2()) %>%
      slice_max(weeks_on_top100, n = 20) %>%
    DT::datatable(colnames = pretty_headers, rownames = FALSE) })

  output$topsong_table <- DT::renderDataTable({
    billboard_dt_new() %>%
        formatStyle("song",
              backgroundColor = "lightblue",
              fontWeight = "bold") %>%
      formatStyle("performer",
              backgroundColor = "pink",
              fontWeight = "bold")
  })
}
  
shinyApp(ui, server)
```

### Question 2. Top Artists

To test that this worked, I searched "Taylor Swift".

```{r topartists_shiny}
### UI ###
ui <- fluidPage(
  title = "Top Artists",
  h2("Top Artists"),
  
  # DROP DOWN
    selectizeInput(
      inputId = "searchme", 
      label = "Search Artist, Select Result, and Chart will Appear",
      multiple = FALSE,
      choices = c("Search Bar" = "", c(unique(billboard_artists$performer))),
      options = list(
        create = FALSE,
        placeholder = "Search Me",
        maxItems = '1',
        onDropdownOpen = I("function($dropdown) {if (!this.lastQuery.length) {this.close(); this.settings.openOnFocus = false;}}"),
        onType = I("function (str) {if (str === \"\") {this.close();}}"))),
  
  #CHART
  highchartOutput("topartist")
  )

### SERVER ###
server <- function(input, output) {

  filtered_billboard_artists <- reactive({
     billboard_artists %>%
      filter(performer == input$searchme) 
  })
  
  output$topartist <- highcharter::renderHighchart({
   filtered_billboard_artists() %>%
    hchart(type = "line", hcaes(x = song_date,
                                y = week_position,
                                group = song)) %>%
      hc_xAxis(title = "Year") %>%
      hc_yAxis(title = "Chart Position",
               reversed = TRUE,
               min = 1,
               max = 100,
               tickmarkPlacement = "on")
    
    })
}
  
shinyApp(ui, server)
```

### Question 3. Audio Features

I chose to visualize the two features of tempo and valence, because I was curious about the relationship between upbeat/happy songs and tempo of it in relation to how successful of a song it is.

```{r spotify_billboards}

### UI ###
ui <- navbarPage(
  
  ## TEXT TITLES##
  title = "Audio Features of Top Songs",
  h2("Audio Features of Top Songs"),
  
  sidebarLayout(
    sidebarPanel(uiOutput("artist_select"),
                 uiOutput("genre_select"),
                 uiOutput("year_select"),
                 uiOutput("top_status_select")),
    mainPanel(plotlyOutput("song_attributes"))))

### SERVER ###
server <- function(input, output) {
  
   ## YEAR INPUT ##
  output$year_select <- renderUI({
    sliderInput(inputId = "choose_year2", 
    label = "Choose the years",
    min = 2004, max = 2014,
    value = c(2004, 2014))})
  
    ## GENRE INPUT ##
  output$genre_select <- renderUI({
    selectInput(inputId = "choose_Genre",
              label = "Genre",
              choices = c("hip hop", "rap", "pop", "rock", "country", "swing",
                          "soul", "blues", "show tunes", "motown", "r&b",
                          "disco", "funk", "rock-and-roll", "freakbeat", "dance"),
              multiple = TRUE,
              selected = c("rap", "pop"))})
  
    ## ARTIST INPUT ##
  output$artist_select <- renderUI({
    selectizeInput(
      inputId = "search_artist2",
      label = "Search Artist, Select Result, and Chart will Appear",
      multiple = FALSE,
      choices = c("Search Bar" = "", c(unique(spotify_billboard$performer))),
      selected = "Katy Perry",
      options = list(
        create = FALSE,
        placeholder = "Search Me",
        maxItems = '1',
        onDropdownOpen = 
          I("function($dropdown) {if (!this.lastQuery.length) {this.close(); this.settings.openOnFocus = false;}}"),
        onType = I("function (str) {if (str === \"\") {this.close();}}")
      ))})
  
   output$top_status_select <- renderUI({
     selectInput(inputId = "choose_TopStatus",
              label = "Top Status",
              choices = c("Top 1" = 1,
                          "Top 10" = 10,
                          "Top 20" = 20),
              multiple = FALSE,
              selected = "Top 1")})
  
    single_artist <- reactive({input$search_artist2})
    
    my_range3 <- reactive({input$choose_year2[1]})
    my_range4 <- reactive({input$choose_year2[2]})
    genre_selection <- reactive({input$choose_Genre})
    
    topstatus <- reactive({input$choose_TopStatus})
    
  
  spotify_billboard_new <- reactive({
    
    spotify_billboard %>%
      dplyr::filter(performer %in% single_artist(),
                    (year >= my_range3() & year <= my_range4()),
                    Top_Type <= topstatus(),
                    str_detect(spotify_genre, paste(genre_selection(), collapse="|"))
                    ) %>%
      rename(Billboard_Peak = Top_Type)
    })
  
  output$song_attributes <- plotly::renderPlotly({
    
      req(spotify_billboard_new())
    
      plot_ly(spotify_billboard_new(),
              x = ~tempo,
              y = ~valence,
              type = "scatter",
              mode = "markers",
              color = ~Billboard_Peak,
              hoverinfo = "text",
              text = ~paste("</br>Artist:", performer,
                            "</br>Song:", song,
                            "</br>Year:", year,
                            "</br>Peak Position:", top_position)) %>%
        layout(title = "Billboard Top Songs' Valence vs Tempo",
         yaxis = list(title = "Valence (0 = sad, 1 = happy)",
                      range = c(0, 1)), 
         xaxis = list(title = "Song Tempo (bpm)",
                      range = c(50, 215)))
  })
    
}
  
shinyApp(ui, server)

```
